---
title: "Pareto Distribution"
output:
  rmarkdown::pdf_document
vignette: >
  %\VignetteIndexEntry{Pareto Distribution}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
bibliography: ["Rrefs.bib"]
biblio-style: "apalike"
link-citations: true
---

```{r echo=FALSE}
library(knitr)
library(pareto)
```

# Introduction

The `pareto` package will compute pdf, cdf and quantile of pareto distribution.

# Pareto distribution
Pareto distribution has density function:
$$f(x|\alpha,\beta) = \frac{\beta\alpha^{\beta}}{x^{\beta+1}}$$
with $0<\alpha\leq{x}$ and $\beta>0$

# compute pdf of pareto distribution

Function "dpareto" will compute the pdf of pareto distribution. It will call dpareto.cpp to complete computation which takes advantage of rcpp to do fast calculation. There're four arguments x,$\alpha$, $\beta$ and log (with default FALSE).The function is fully vectorized.

There's an simple example with $x=c(3,4)$, $\alpha = c(2,3)$, $\beta=2$
```{r}
dpareto(c(3,4),c(2,3),2)
```

When either $\alpha$ or $\beta$ is less or equal to 0, function dpareto will return NaN.

```{r}
dpareto(3,-1,2)
dpareto(3,1,0)
```

In addition, when x is less than $\alpha$, the pareto density is obviously 0.

```{r}
dpareto(1.9,2,2)
```

Furthermore, we can ask function dpareto to return log density with argument "log = T"

```{r}
dpareto(3,2,2, log = T)
```

An example plot of pareto density, with $\alpha=2$ and $\beta=2$
```{r}
x <- seq(2,20,length.out = 1000)

plot(x,dpareto(x,alpha=2,beta=2), main = "pareto density(alpha = beta = 2)", ylab = "density")
```

# compute cdf of pareto distribution

Function "ppareto" will compute the cdf of pareto distribution. There're three arguments x,$\alpha$ and $\beta$. The function is fully vectorized.

There's an simple example with $x=c(3,4)$, $\alpha = c(2,3)$, $\beta=2$
```{r}
ppareto(c(3,4),c(2,3),2)
```

Also, when either $\alpha$ or $\beta$ is less or equal to 0, function ppareto will return NaN.

```{r}
ppareto(3,-1,2)
ppareto(3,1,0)
```

# compute quantile of pareto distribution

Function "qpareto" will compute the quantile of pareto distribution. There're three arguments p,$\alpha$ and $\beta$. The function is fully vectorized.

```{r, warning=FALSE}
qpareto(c(0.3,0.4,0.5),2,3)
```

Also, when either $\alpha$ or $\beta$ is less or equal to 0, function qpareto will return NaN.

```{r}
qpareto(0.3,-1,2)
qpareto(0.3,1,0)
```

A table of pareto ($\alpha=2$,$\beta=2$) pdf and cdf

```{r}
x.table <- seq(3,10,by = 1)

p.pdf <- dpareto(x.table,alpha = 2,beta = 2)

p.cdf <- ppareto(x.table,alpha = 2,beta = 2)

p.table <- data.frame(x.table, p.pdf, p.cdf)

names(p.table) <- c("x", "pdf", "cdf")

kable(p.table)
```


# Acknowledgement

This document depended on the packages

* markdown [@R-markdown]
* knitr [@R-knitr]

# References

* Pareto [@Pareto]