---
title: "numderivs"
author: "Ruifeng Chen"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
h <- NULL

for (i in 1:64){
  h[i] <- 2^(-i)
}
```

Approximations to the derivative of a function f. There are three methods to get the derivarive.

(1) forward difference quotient (for a step size h):
$$\partial_F(f,x,h) = \frac{f(x+h)-f(x)}{h}$$

(2) symmetric difference quotient (for a step size h):
$$\partial_S(f,x,h) = \frac{f(x+h)-f(x-h)}{2h}$$
(3) when the function f is analytic near x:
$$\partial_C(f,x,h) = \frac{\Im(f(x+hi))}{h}$$
where $i = \sqrt{-1}$ is the imaginary unit and $\Im(z)$ is the imaginary part of the complex number z.

We choose a few functions and argument values and examine these approximations graphically. The y axis will denote the derivative, and x axis will denote $-log_2h$, where h is shown above.

The first example we use is $f_1(x) = sin(x)$ at $x = 1$.
```{r}
x <- 1
derivative1 <- NULL
derivative2 <- NULL
derivative3 <- NULL
```

```{r}
#derivative1
for(i in 1:64){
  derivative1[i] <- (sin(x + h[i]) - sin(x)) / h[i]
}
plot(seq(1:64), derivative1, main = "forward difference quotient")
```

```{r}
#derivative2
for(i in 1:64){
  derivative2[i] <- (sin(x + h[i]) - sin(x - h[i])) / (2*h[i])
}
plot(seq(1:64), derivative2, main = "symmetric difference quotient")
```

```{r}
#derivative3
for(i in 1:64){
  derivative3[i] <- Im(sin(complex(real = x, imaginary = h[i])))/h[i]
}
plot(seq(1:64), derivative3, main = "the third derivative method")
```

In the first example, we can see that derivatives derived by three methods respectively are quite stable among h values $2^{10}$ to $2^{40}$. However, they are unstable with very small or very large h values.


The second example we use is $f_2(x) = 10000sin(x)$ at $x = 1$.
```{r}
x <- 1
derivative1 <- NULL
derivative2 <- NULL
derivative3 <- NULL
```

```{r}
for(i in 1:64){
  derivative1[i] <- 10000 * (sin(x + h[i]) - sin(x)) / h[i]
}
plot(seq(1:64), derivative1, main = "forward difference quotient")
```

```{r}
for(i in 1:64){
  derivative2[i] <- 10000 * (sin(x + h[i]) - sin(x - h[i])) / (2*h[i])
}
plot(seq(1:64), derivative2, main = "symmetric difference quotient")
```

```{r}
for(i in 1:64){
  derivative3[i] <- Im(10000*sin(complex(real = x, imaginary = h[i])))/h[i]
}
plot(seq(1:64), derivative3, main = "the third derivative method")
```

In the second example, the result is very similar to the first example.

The third example we use is $f_3(x) = tan(x)$ at $x = 1.59$.
```{r}
x <- 1.59
derivative1 <- NULL
derivative2 <- NULL
derivative3 <- NULL
```

```{r}
for(i in 1:64){
  derivative1[i] <- (tan(x + h[i]) - tan(x)) / h[i]
}
plot(seq(1:64), derivative1, main = "forward difference quotient")
```

```{r}
for(i in 1:64){
  derivative2[i] <- (tan(x + h[i]) - tan(x - h[i])) / (2*h[i])
}
plot(seq(1:64), derivative2, main = "symmetric difference quotient")
```

```{r}
for(i in 1:64){
  derivative3[i] <- Im(tan(complex(real = x, imaginary = h[i]))) / h[i]
}
plot(seq(1:64), derivative3, main = "the third derivative method")
```

In the third example, we can see that derivatives derived by three methods respectively are quite stable among h values $2^{10}$ to $2^{50}$. However, they are unstable with very small or very large h values.

In all the three examples, we can see that these derivatives are quite stable in the middle range, but unstable with extreme values. This is due to floating
point arithmetic.

This suggests that in practice, we can choose size h among $2^{-20}$ to $2^{-40}$ to get pretty stable derivatives.

