---
title: "pareto-intro"
author: "Ruifeng Chen"
date: "2018/10/17"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The package dpareto could calculate the density of pareto distribution. The main function of dpareto package is function dpareto, which would call Rcpp to calculate density of pareto distribution in a fast way. 
```{r}
library(pareto)
```

In total, there's 4 arguments for function dpareto, they are x, $\alpha$, $\beta$ and log. The density of pareto distribution is:
$$f(x|\alpha,\beta) = \frac{\beta\alpha^{\beta}}{x^{\beta+1}}$$
The log argument is TRUE or FALSE, which indicates whether we want a log density or not.

The function is fully vectorized thus all of x, $\alpha$ and $\beta$ could be vectors. Obviously, it means we can get multi-densities simultaneously.

There's an simple example with $x=3$, $\alpha = 2$, $\beta=2$
```{r}
dpareto(x = 3,alpha = 2,beta = 2,log = FALSE)
```

There's another example with vectorized x, $\alpha$ and $\beta$
```{r}
dpareto(x = c(3,4,5),alpha = c(2,3),beta = 2,log = FALSE)
```

In addition, we if any of $\alpha$ or $\beta$ is negative or NaN, the density would be NaN
```{r}
dpareto(x = c(3,4,5),alpha = c(-2,3),beta = 2,log = FALSE)
dpareto(x = c(3,4,5),alpha = c(2,3),beta = -2,log = FALSE)
```

Furthermore, when the x is less than $\alpha$, the density should be 0.
```{r}
dpareto(x = c(3,3,5),alpha = c(2,4),beta = 2,log = FALSE)
```

